Boosted Regression Trees for Predictive Auto-Tuning

### 2011-12-13 paper/draft1.tex
Minor edits to the abstract

\begin{abstract}
Auto-tuning of parameteric, instrumented code templates has proven to be a highly effective technique for optimizing a GPU kernel for a particular computation on particular
hardware \cite{XXXX}.  
% DDC: I'm still struggling with the above sentence.  Feel free to substitute your own ideas
However, a drawback of this approach is that exhaustive, or even thorough, auto-tuning requires
compiling many kernels and calling each one many times, and this process is slow.
Furthermore, it is desirable to provide users with unified library abstraction boundaries for operations such as image
filtering and matrix multiplication, even those these operations actually correspond to a large set
of potential problem configurations with a wide variety of memory access
patterns and computational bottlenecks.  How can we draw on data from previous
empirical auto-tuning of related problems on related hardware to make a
just-in-time implementation decision for a novel problem?  This paper presents
a machine learning approach to auto-tuning, in which features of the
current hardware platform, the kernel configuration and the problem
instance are passed to a regression model (boosted regression trees) which
predicts how much faster this kernel will be than a reference baseline.
Combinatorial optimization strategies for auto-tuning that would normally 
require evaluating large number of kernel configurations on the real hardware, are made orders of
magnitude faster by evaluating the surrogate regression model instead.  We
validate our approach using the filterbank correlation kernel described in
\citet{pinto+cox:2011gcg}, where we find that 0.1 seconds (XXX) of hill
climbing on the regression model (which we dub ``{\em predictive} auto-tuning'') can achieve an
average of 95\% of the speed-up brought by 120 seconds of empirical auto-tuning
XXX.  Our approach is not specific to filterbank correlation, nor even to GPU
kernel auto-tuning, and can be applied to almost any templated-code optimization problem, spanning a wide variety of problem types, kernel types, and
platforms.  \end{abstract}

>>>>

\begin{abstract}
Auto-tuning of parameteric, instrumented code templates has proven to be a
highly effective technique for optimizing a GPU kernel for a particular
computation on particular
hardware \cite{XXXX}.  
% DDC: I'm still struggling with the above sentence.  Feel free to substitute your own ideas
However, a drawback of this approach is that exhaustive, or even thorough, auto-tuning requires
compiling many kernels and calling each one many times, and this process is slow.
Furthermore, it is desirable to provide users with unified library abstraction
boundaries for operations such as image
filtering and matrix multiplication, even those these operations actually correspond to a large set
of potential problem configurations with a wide variety of memory access
patterns and computational bottlenecks.  How can we draw on data from previous
empirical auto-tuning of related problems on related hardware to make a
just-in-time implementation decision for a novel problem?  This paper presents
a machine learning approach to auto-tuning, in which features of the
current hardware platform, the kernel configuration and the problem
instance are passed to a regression model (boosted regression trees) which
predicts how much faster this kernel will be than a reference baseline.
Combinatorial optimization strategies for auto-tuning that would normally
require evaluating large number of kernel configurations on the real hardware, are made orders of
magnitude faster by evaluating the surrogate regression model instead.  We
validate our approach using the filterbank correlation kernel described in
\citet{pinto+cox:2011gcg}, where we find that 0.1 seconds (XXX) of hill
climbing on the regression model (which we dub ``{\em predictive} auto-tuning'') can achieve an
average of 95\% of the speed-up brought by minutes of empirical auto-tuning
XXX.  Our approach is not specific to filterbank correlation, nor even to GPU
kernel auto-tuning, and can be applied to almost any templated-code
optimization problem, spanning a wide variety of problem types, kernel types,
and platforms.

\end{abstract}
